---
title: "Hospital Length of Stays"
author: "Keith Douglas"
date: "2024-10-15"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
library(NHSRdatasets)
library(knitr)
library(kableExtra)
```

# Load the data from the package
```{r}
data("LOS_model")
?LOS_model
```

# Inspect
```{r}
summary(LOS_model)

```


# Make Death a factor
```{r}
LOS_model$Death = as.factor(LOS_model$Death)
```


# Recode Death levels
```{r}
mut_df = LOS_model %>% 
  mutate(Death = Death %>% 
           fct_recode("Survived" = "0", "Died" = "1"))
```


# Create a summary table where each combination of Organisation and Death gets a count (n).
```{r}
# table(mut_df$Organisation, mut_df$Death)

summ_df = mut_df %>% 
  group_by(Organisation, Death) %>% 
  tally()
```

# Make a wide table with Dead and Survived as rows with a column for each Trust
```{r}
# t(table(mut_df$Organisation, mut_df$Death))

wide_df = summ_df %>% 
  pivot_wider(names_from = Organisation,
               values_from = n)
```

# Another pivot with Survived and Died as columns, Trusts as rows.
# Also calculate the % survived for each Trust
```{r}
perc_df = summ_df %>% 
  pivot_wider(
    names_from = Death,
    values_from = n
  ) %>% 
  mutate(Total = Survived + Died,
         Percent_Survived = (Survived / Total) * 100) %>% 
  mutate(Organisation = Organisation %>% 
           fct_recode("Trust 1" = "Trust1", "Trust 2" = "Trust2", "Trust 3" = "Trust3", "Trust 4" = "Trust4", "Trust 5" = "Trust5", "Trust 6" = "Trust6", "Trust 7" = "Trust7", "Trust 8" = "Trust8", "Trust 9" = "Trust9", "Trust 10" = "Trust10"))
```


# Make the wide table pretty with kable()
```{r}

perc_df %>% 
  kable(
    col.names = c("Trust", "Survived", "Died", "Total", "Survived (%)"),
    digits = 0, # Rounding up stuff
    caption = "Hospital Length of Stays data: Percent Survived by Trust",
    align = "lcccc"
    ) #%>% 
  #kable_styling(
   # "striped",
    #full_width = FALSE
  #) %>% 
  #footnote("Data from LOS_model")
# kableExtra stuff usually doesn't knit well to pdf
```


# Let's knit to PDF


